---
title: "Greenhouse and Molecular correlation"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



```{r load packages, include=FALSE}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(ggrepel)
library(gganimate)
```


# Molecular and greenhouse validation of field evolved resistance to glyphosate and PPO-inhibitors in Palmer amaranth populations

Methods: Glyphosate was applied 


```{r load full data, include=FALSE}
Data=read_csv("Survival.csv")
glimpse(Data)
Data$Time=factor(Data$Time, levels=c("Winter", "Summer", "Fall"), labels=c("Winter", "Summer", "Fall"))
```


# Gliphosate Resistant Population

```{r filter Gly and R, include=FALSE}
DataRG=filter(Data, Herbicide=="Glyphosate" & Level=="Resistant")
```


```{r correlation, echo=FALSE}
DataRGS=filter(DataRG, Time=="Winter")
cor.test(DataRGS$Perc, DataRGS$PercMolec, method = "pearson", conf.level = 0.95)
DataRGSu=filter(DataRG, Time=="Summer")
cor.test(DataRGSu$Perc, DataRGSu$PercMolec, method = "pearson", conf.level = 0.95)
```


```{r, Figure GLY and Resistant, fig.width = 12, fig.height= 6}
ggplot(DataRG, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") +
  facet_grid(~Herbicide) + labs(x="% of survival plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), 
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 23, y = 100, label = c("Pearson's correlation = 0.43 (P=0.06)", "Pearson's correlation = 0.86 (P=0.00)")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + facet_wrap(~Time) +
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("GlyphosateR.tiff", units="in", width=12, height=6, dpi=600) 
```





# Gliphosate Susceptible Population


```{r filter GLY and S, echo=FALSE, include=FALSE}
DataSG=filter(Data, Herbicide=="Glyphosate" & Level=="Susceptible")
```

```{r correlation GLY and S, filter GLY and S, echo=FALSE}
DataSGS=filter(DataSG, Time=="Winter")
cor.test(DataSGS$Perc, DataSGS$PercMolec, method = "pearson", conf.level = 0.95)
DataSGSu=filter(DataSG, Time=="Summer")
cor.test(DataSGSu$Perc, DataSGSu$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, Figure GLY and Susceptible, fig.width = 12, fig.height= 6, echo=FALSE}
ggplot(DataSG, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") +
  facet_grid(~Herbicide) + labs(x="% of dead plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), 
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 75, y = 100, label = c("Pearson's correlation = -0.27 (P=0.25)", "Pearson's correlation = -0.86 (P=0.00)")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + facet_wrap(~Time) +
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("GlyphosateS.tiff", units="in", width=12, height=6, dpi=300) 
```



# Gliphosate Medium Population

```{r load Gly and M, echo=FALSE, include=FALSE}
DataMG=filter(Data, Herbicide=="Glyphosate" & Level=="Medium")
```


```{r Correlation Gly and M, echo=FALSE}
DataMGS=filter(DataMG, Time=="Winter")
cor.test(DataMGS$Perc, DataMGS$PercMolec, method = "pearson", conf.level = 0.95)
DataMGSu=filter(DataMG, Time=="Summer")
cor.test(DataMGSu$Perc, DataMGSu$PercMolec, method = "pearson", conf.level = 0.95)
```

```{r, fig.width = 12, fig.height= 6, echo=FALSE}
ggplot(DataMG, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") +
  facet_grid(~Herbicide) + labs(x="% of medium injury plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), 
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 78, y = -15, label = c("Pearson's correlation = -0.06 (P=0.81)", "Pearson's correlation = 0.87 (P=0.00)")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + facet_wrap(~Time) +
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("GlyphosateM.tiff", units="in", width=12, height=6, dpi=300) 
```




# Glyphosate High Population

```{r include=FALSE}
DataLG=filter(Data, Herbicide=="Glyphosate" & Level=="Low")
```


```{r}
DataLGS=filter(DataLG, Time=="Winter")
cor.test(DataLGS$Perc, DataLGS$PercMolec, method = "pearson", conf.level = 0.95)
DataLGSu=filter(DataMG, Time=="Summer")
cor.test(DataLGSu$Perc, DataLGSu$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, fig.width = 12, fig.height= 6, echo=FALSE}
ggplot(DataLG, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") +
  labs(x="% of low injury plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), 
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 85, y = -17, label = c("Pearson's correlation = 0.21", "Pearson's correlation = 0.87")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + facet_grid(~Time) +
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("GlyphosateL.tiff", units="in", width=12, height=6, dpi=300) 
```







# Fomesafen Resistant


```{r}
DataRF=filter(Data, Herbicide=="Fomesafen" & Level=="Resistant")
DataRF$Time=factor(DataRF$Time, levels=c("Winter", "Summer", "Fall"), labels=c("Winter", "Summer", "Fall"))
```


```{r}
FomesafenRS=subset(DataRF, Time=="Winter")
cor.test(FomesafenRS$Perc, FomesafenRS$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenRSu=subset(DataRF, Time=="Summer")
cor.test(FomesafenRSu$Perc, FomesafenRSu$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenRF=subset(DataRF, Time=="Fall")
cor.test(FomesafenRF$Perc, FomesafenRF$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(DataRF, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") + facet_grid(~Time) +
  labs(x="% of survival plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = 0.14", "Pearson's correlation = 0.23", "Pearson's correlation = 0.40")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("FomesafenR.tiff", units="in", width=12, height=6, dpi=300) 
```




# Fomesafen Susceptible


```{r}
DataSF=filter(Data, Herbicide=="Fomesafen" & Level=="Susceptible")
DataSF$Time=factor(DataSF$Time, levels=c("Spring", "Summer", "Fall"), labels=c("Spring", "Summer", "Fall"))
```


```{r}
FomesafenSS=subset(DataSF, Time=="Spring")
cor.test(FomesafenSS$Perc, FomesafenSS$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenSSu=subset(DataSF, Time=="Summer")
cor.test(FomesafenSSu$Perc, FomesafenSSu$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenSF=subset(DataSF, Time=="Fall")
cor.test(FomesafenSF$Perc, FomesafenSF$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(DataSF, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") + facet_grid(~Time) +
  labs(x="% of dead plants 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = -0.14", "Pearson's correlation = -0.22", "Pearson's correlation = -0.40")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("FomesafenS.tiff", units="in", width=12, height=6, dpi=300) 
```


# Fomesafen Medium


```{r}
DataMF=filter(Data, Herbicide=="Fomesafen" & Level=="Medium")
DataMF$Time=factor(DataMF$Time, levels=c("Spring", "Summer", "Fall"), labels=c("Spring", "Summer", "Fall"))
```


```{r}
FomesafenMS=subset(DataMF, Time=="Spring")
cor.test(FomesafenMS$Perc, FomesafenMS$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenMSu=subset(DataMF, Time=="Summer")
cor.test(FomesafenMSu$Perc, FomesafenMSu$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenMF=subset(DataMF, Time=="Fall")
cor.test(FomesafenMF$Perc, FomesafenMF$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(DataMF, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") + facet_grid(~Time) +
  labs(x="% of plants with medium injury 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = 0.18", "Pearson's correlation = 0.35", "Pearson's correlation = 0.40")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("FomesafenM.tiff", units="in", width=12, height=6, dpi=300) 
```




# Fomesafen Low


```{r}
DataLF=filter(Data, Herbicide=="Fomesafen" & Level=="Low")
DataLF$Time=factor(DataLF$Time, levels=c("Spring", "Summer", "Fall"), labels=c("Spring", "Summer", "Fall"))
```


```{r}
FomesafenLS=subset(DataLF, Time=="Spring")
cor.test(FomesafenLS$Perc, FomesafenLS$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenLSu=subset(DataLF, Time=="Summer")
cor.test(FomesafenLSu$Perc, FomesafenLSu$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenLF=subset(DataLF, Time=="Fall")
cor.test(FomesafenLF$Perc, FomesafenLF$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(DataLF, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") + facet_grid(~Time) +
  labs(x="% of plants with low injury 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = -0.31", "Pearson's correlation = -0.33", "Pearson's correlation = 0.40")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("FomesafenL.tiff", units="in", width=12, height=6, dpi=300) 
```



# Lactofen

```{r}
DataML=filter(Data, Herbicide=="Lactofen" & Level=="Resistant")
DataML$Time=factor(DataML$Time, levels=c("Spring", "Summer", "Fall"), labels=c("Spring", "Summer", "Fall"))
```


```{r}
LactofenMS=subset(DataMF, Time=="Spring")
cor.test(FomesafenMS$Perc, FomesafenMS$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenMSu=subset(DataMF, Time=="Summer")
cor.test(FomesafenMSu$Perc, FomesafenMSu$PercMolec, method = "pearson", conf.level = 0.95)

FomesafenMF=subset(DataMF, Time=="Fall")
cor.test(FomesafenMF$Perc, FomesafenMF$PercMolec, method = "pearson", conf.level = 0.95)
```



```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(DataMF, aes(x=Perc, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") + facet_grid(~Time) +
  labs(x="% of plants with medium injury 21 days after herbicide application", y="% resistant plants in a molecular screening") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = 0.18", "Pearson's correlation = 0.35", "Pearson's correlation = 0.40")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("FomesafenM.tiff", units="in", width=12, height=6, dpi=300) 
```






```{r}
Corr=read_csv("PPO_COR.csv")
```


```{r}
cor.test(Corr$PercSurvL, Corr$PercMolec, method = "pearson", conf.level = 0.95)


cor.test(Corr$PercSurvF, Corr$PercMolec, method = "pearson", conf.level = 0.95)


cor.test(Corr$PercSurvF, Corr$PercSurvL, method = "pearson", conf.level = 0.95)
```


```{r, echo=FALSE, fig.width = 12, fig.height= 12}
ggplot(Corr, aes(x=PercSurvL, y=PercMolec, label=Population)) + geom_point(aes(size=AvegGA, color=PPO)) + geom_smooth(method='lm',formula=y~x, se=F, color="blue") +
  labs(x="% of Lactofen", y="% Molecular") + 
  theme(axis.text=element_text(size=11, color="black"), 
        axis.title=element_text(size=13,face="bold"), legend.background = element_rect(linetype = "solid", color="black"), #panel.border = element_rect(colour = "black", fill=NA, size=5),
        panel.background = element_rect(fill="white", color = "black"), legend.position=c("bottom"), legend.title = element_text(face="bold"), legend.key = element_blank(), 
        panel.grid.major = element_line(color = "white"), plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
  panel.grid.minor = element_blank(), strip.background = element_rect(fill="grey95", color="white"), strip.text.y = element_text(size = 16, color="white"), strip.text.x = element_text(size = 16, color="black")) + 
  geom_label_repel(aes(label = Population),
                box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50', size=3) + annotate("text", x = 60, y = -17, label = c("Pearson's correlation = 0.14")) +
  scale_colour_manual(name="PPO resistance:", values = c("darkgreen", "red")) + 
   scale_size_continuous(name="EPSPS gene amplification:", range=c(1,5), limits=c(1,50), breaks=c(1,10,20,30,40),labels=c("<2","3 to 10","11 to 20","21 to 30",">30")) + 
    ggsave("PPOcorr.tiff", units="in", width=6, height=6, dpi=300) 
```




