% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  a4paper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Distribution and Validation of EPSPS- and PPO-Inhibitor Resistance in Amaranthus palmeri from Southwestern Nebraska.},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{float}
\usepackage{placeins}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{lineno}
\usepackage{setspace}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}

\title{Distribution and Validation of EPSPS- and PPO-Inhibitor Resistance in
\emph{Amaranthus palmeri} from Southwestern Nebraska.}
\author{true \and true \and true \and true \and true}
\date{}

\begin{document}
\maketitle

\vspace{2mm}

\doublespace

\large Abstract

\textbf{BACKGROUNG:} Failure to control \textit{Amaranthus palmeri} with
5-enolpyruvylshikimate-3-phosphate synthase (EPSPS)- and
protoporphyrinogen oxidase (PPO)-inhibitor herbicides was reported
across southwestern Nebraska in 2017. The objectives of this study were
to (1) confirm and validate EPSPS (glyphosate)- and PPO (fomesafen and
lactofen)-resistance in 51 \textit{A. palmeri} populations from
southwestern Nebraska using genotypic and whole-plant phenotypic assays,
and (2) determine which agronomic practices are influencing
EPSPS-inhibitor resistance in \textit{A. palmeri} populations in that
geography.\\
\textbf{RESULTS:} Based on genotypic assay, 88\% of 51 populations
contained at least one individual with \textit{EPSPS} gene amplification
(\textgreater{} 2 copies), which confers glyphosate resistance; or a
mutation in the \textit{PPX2} gene \(\triangle\)G210 or R128G/M,
endowing PPO-inhibitor resistance in \textit{A. palmeri}. High
correlation (0.83) between genotypic and phenotypic assays demonstrated
that \textit{EPSPS} gene amplification is the main glyphosate resistance
mechanism in \textit{A. palmeri} populations from southwesten Nebraska.
In contrast, there was poor association between genotypic and phenotypic
responses for PPO resistance, which we attribute to a combination of
other resistance mechanisms being present, segregation of the
populations, and using too high of herbicide doses. Thus, genotypic
assays could expedite the process for confirmation of EPSPS- but not
necessarily for PPO-inhibitor resistance in \textit{A. palmeri} from
southwestern Nebraska. Moreover, \textit{EPSPS} gene amplification,
county, current and previous crops are the main factors influencing
glyphosate resistance within that geography. Most of EPSPS-susceptible
\textit{A. palmeri} populations were found in a few counties in high
crop diversity areas.\\
\textbf{CONCLUSIONS:} Results presented here confirmed the spread of
EPSPS- and PPO-inhibitor resistance in \textit{A. palmeri} populations
from southwestern Nebraska and show that less diverse cropping systems
are an important driver for EPSPS-inhibitor resistance evolution in
\textit{A. palmeri}.

\vspace{2mm}

\noindent \textbf{Keywords}: Glyphosate; Fomesafen; Lactofen; Palmer
amaranth; Weed resistance.

\linenumbers

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\emph{Amaranthus palmeri} S. Watson (Palmer amaranth) is a plant species
indigenous to the southwestern United States (US) and northern Mexico
(Sauer 1957). Despite being an edible plant used to feed humans and
animals (Smith 1900), \emph{A. palmeri} has long been documented as a
serious weed problem in US cropping systems (Hamilton and Arle 1958).
\emph{A. palmeri} produces thousands of seeds and grows up to 2 m tall
with many lateral branches (Sauer 1955), making it a very competitive
species with crops (Massinga et al. 2003, Morgan et al. 2001). In the
1970s, when cotton picking became mechanized, machinery contributed to
the spread of \emph{A. palmeri} seeds across the southern United States
(Sauer 1972). At that time, \emph{A. palmeri} was considered the most
successful weed of all dioecious \emph{Amaranthus} species as it became
widespread in cotton fields (Sauer 1972). Currently, \emph{A. palmeri}
is the most economically damaging weed species infesting corn, cotton,
and soybean fields in the southern US (Price et al. 2011, Ward et al.
2013).

obligate cross-pollination in A. palmeri is herbicide resistance
movement via gene flow within and among populations within the species

The economic importance of \emph{A. palmeri} is primarily related to its
ability to evolve resistance to herbicides. Factors relating to the
intrinsic biology of this species have contributed to its fast herbicide
resistance evolution (Ward et al. 2013). For example, \emph{A. palmeri}
reproduces via obligate cross pollination, increasing the chances of
herbicide resistance alleles transfer via gene flow within and among
populations (Gaines et al. 2012, Oliveira et al. 2018). In addition,
human-driven selection has strongly contributed to the rise of \emph{A.
palmeri} as a problematic weed. In the US cropping systems, no-till is a
widely used practice amongst growers, and \emph{A. palmeri} thrives in
no-till fields due to its small seed size, contributing to the rapid
increase of \emph{A. palmeri} individuals in field crops. Also,
herbicide resistance in \emph{A. palmeri} drastically increased when
weed management strategies shifted from the use of multiple herbicide
sites of action (SOA) in a single season to reliance on a single SOA
post-emergence herbicide (e.g., glyphosate) (Powles 2008). Thus far,
\emph{A. palmeri} has evolved resistance to herbicide resistance to
multiple SOA (Heap 2019), which is a major concern as weed management in
conventional US cropping systems is largely herbicide dependent.

The history of herbicide resistance evolution in \emph{A. palmeri} is a
result of intense selection pressure from herbicides. In the 1990s, the
first documented cases of herbicide resistance were against microtubule-
(Gossett et al. 1992), ALS- (Horak and Peterson 1995), and
PSII-inhibitor herbicides (Heap 2019). With the introduction of
glyphosate-resistant (GR) crops, post-emergence applications of
glyphosate became widely used for weed management in soybean, cotton,
and corn, resulting in evolution of GR \emph{A. palmeri} (Culpepper et
al. 2006). GR \emph{A. palmeri} spread across the southern and
midwestern US through both independent herbicide selection (Küpper et
al. 2018) and seed dispersal (Farmer et al. 2017, Norsworthy et al.
2014). The spread of GR \emph{A. palmeri} led to a reevaluation of the
use of glyphosate as a sole means of weed control in this species, and a
push to diversify weed management strategies (e.g., application of other
herbicide SOA). The use of 4-hydroxyphenylpyruvate dioxygenase (HPPD)-,
protoporphyrinogen oxidase (PPO)- and long chain fatty acids (LCFA)-
inhibitor herbicides increased in an attempt to manage GR \emph{A.
palmeri}. However, \emph{A. palmeri} populations also evolved resistance
to HPPD-, PPO- and LCFA-inhibitor herbicides (Heap 2019). New
technologies such as auxin-resistant crops may be jeopardized by the
newest reports of 2,4-D-resistant \emph{A. palmeri} (Heap 2019) and the
number of \emph{A. palmeri} populations with herbicide resistance to
multiple herbicide resistance is also on the rise (Kumar et al. 2019,
Schwartz-Lazaro et al. 2017). Therefore, \emph{A. palmeri} herbicide
resistance evolution is shrinking the chemical control options for weed
management in corn, soybean, and cotton fields within US cropping
systems.

In the US Midwest, corn and soybean growers strongly rely on EPSPS-
(e.g., glyphosate) and PPO- (e.g., fomesafen) inhibitor herbicides for
weed management. The recent migration of \emph{A. palmeri} into the US
Midwest poses a serious threat to the sustainability of crop production
in that geography (Chahal et al. 2017, Kohrt et al. 2017). \emph{A.
palmeri} is now overlapping territory with another problematic dioecious
\emph{Amaranthus} species, waterhemp (\emph{Amaranthus tuberculatus}).
Thus, prevention and/or rapid diagnosis of herbicide resistance in
\emph{A. palmeri} has become a priority for agricultural stakeholders.
The advances in high-throughput genome sequencing methods are expediting
the elucidation and detection of herbicide resistance mechanisms in
\emph{A. palmeri} and other weed species. In the case of glyphosate, the
most common resistance mechanism in \emph{A. palmeri} is \emph{EPSPS}
gene amplification (Gaines et al. 2019, 2011), while in PPO-inhibitor
resistance, the major resistance mechanism is a \emph{PPO2} glycine 210
deletion (\(\triangle\)G210) (Salas et al. 2016, Salas-Perez et al.
2017). Nonetheless, novel herbicide resistance mechanisms in \emph{A.
palmeri} are still being uncovered (Gaines et al. 2019), as evidenced by
the recent documentation of two mutations in the \emph{PPO2} enzyme in
the R128 site of \emph{A. palmeri} (Giacomini et al. 2017), and G399A,
an amino acid substitution of glycine to alanine in the catalytic domain
of \emph{PPO2} at position 399 (Rangani et al. 2019). So far, a few
reports of resistance mechanisms have been confirmed for \emph{A.
palmeri} populations with non-target-site resistance to glyphosate
(Dominguez-Valenzuela et al. 2017, Palma-Bautista et al. 2019) and one
against PPO-inhibitor herbicides (Varanasi et al. 2018a). Therefore,
using genotypic assays might provide faster detection of known herbicide
weed resistance mechanisms in \emph{A. palmeri}, but it fails to address
herbicide resistance resulting from novel mechanisms, including
polygenic resistance mechanisms (e.g.~metabolic resistance involving
cytochrome P450 genes) as well as previously unknown target site
mutations.

In the fall of 2017, growers in southwestern Nebraska reported failure
to control \emph{A. palmeri} with EPSPS- and PPO-inhibitor herbicides
(Werle R, personal communication). The objectives of this study were to
(1) confirm EPSPS- and PPO-inhibitor resistance in 51 \emph{A. palmeri}
populations from southwestern Nebraska via genotypic resistance assays
and compare these results via whole-plant phenotypic assay of their
progenies, and (2) evaluate agronomic practices that may contribute to
EPSPS- and PPO-inhibitor resistance in \emph{A. palmeri} populations.

\hypertarget{material-and-methods}{%
\section{Material and Methods}\label{material-and-methods}}

\hypertarget{plant-material-and-growing-conditions}{%
\subsection{Plant Material and Growing
Conditions}\label{plant-material-and-growing-conditions}}

The study was performed with 51 arbitrarily selected \emph{A. palmeri}
populations infesting cropping systems across Southwestern Nebraska.
Each population was collected from a single field. Location, agronomic
practices, \emph{A. palmeri} distribution and density of each population
were collected (Table 1). In August 2017, green leaf tissues were
harvested from five actively growing plants from each of the 51 \emph{A.
palmeri} populations, then labeled and stored at -80 C to be used in
genotypic assays. Within the 51 \emph{A. palmeri} populations, a second
sampling of 19 arbitrarily selected populations was obtained by
collecting seeds (progenies) from 30 plants each in September 2017, then
cleaned, bulked for each population, and stored at 5 C until the onset
of the greenhouse experiments. Seeds were planted in 900
cm\textsuperscript{-3} plastic trays containing potting-mix
(Pro-Mix\textsuperscript \textregistered, Quakertown, PA, USA). Emerged
seedlings (1 cm) were transplanted into 164 cm\textsuperscript{-3}
cone-tainers. \emph{A. palmeri} plants were supplied with adequate water
and kept under greenhouse conditions at 28/20 C day/night temperature
with 80\% relative humidity. Artificial lighting was provided using
metal halide lamps (600 \(\mu\)mol m\textsuperscript{-2}
s\textsuperscript{-1}) to ensure 15 h photoperiod.

\hypertarget{genotypic-herbicide-resistance-mechanism-assays}{%
\subsection{Genotypic Herbicide Resistance Mechanism
Assays}\label{genotypic-herbicide-resistance-mechanism-assays}}

As a standard methodology from the University of Illinois Plant Clinic
Laboratory (Urbana,85IL 61801, United States), three to five leaf tissue
samples were collected from 51 \emph{A. palmeri} populations from
Southwestern Nebraska. The EPSPS copy number was estimated for each
plant based on DNA extracted from tissue from a single leaf of \emph{A.
palmeiri}. Genomic DNA extraction from leaf tissue87samples were
performed using a modified CTAB method (Doyle and Doyle 1987). DNA
quality and quantity were checked on a Nanodrop 1000 (Thermo Fisher
Scientific, Inc., Waltham, MA, USA) and any samples with low DNA yields
or high protein:DNA ratios were discarded and re-extracted. TaqMan qPCR
assays were used to check for the presence of known PPO-inhibitor
resistance mutations on the \emph{PPO2} enzyme, including the glycine
210 deletion (Wuerffel et al. 2015) and the R128G/M mutations (Varanasi
et al. 2018b). Samples were also tested for glyphosate resistance via
increased numbers of \emph{EPSPS} genomic copies using a SYBR qPCR
approach (Chatham et al. 2015) in which \emph{EPSPS} copy numbers were
estimated based on comparison with a single-copy reference gene
(\emph{CPS}, carbamoyl phosphate synthetase).

Herein, for the objective of this study, individuals possessing
\emph{EPSPS} copy numbers \textgreater{} 2 are considered
EPSPS-resistant and individuals with presence of \(\triangle\)G210 or
R128 het mutations are considered PPO-resistant. Other target-site and
non-target site resistance mechanisms were not tested.

\hypertarget{whole-plant-phenotypic-assay}{%
\subsection{Whole-plant Phenotypic
Assay}\label{whole-plant-phenotypic-assay}}

The research was conducted under greenhouse conditions in 2018 and 2019
at the University of Wisconsin-Madison to evaluate the sensitivity of 19
\emph{A. palmeri} populations from Southwestern Nebraska to EPSPS- and
PPO-inhibitor herbicides.

The experiments were conducted in a complete randomized design and the
experimental unit was a cone-tainer with a single \emph{A. palmeri}
seedling. The study was arranged in a factorial design with \emph{A.
palmeri} progenies from 19 populations and three herbicides with 20
replications and conducted twice. Altogether, 2280 \emph{A. palmeri}
seedlings were screened in the phenotypic assay. The arbitrarily
selected 19 \emph{A. palmeri} progenies were from Cha 3, Dun 3, Dun 4,
Dun 5, Hay 1, Hay 3, Hay 4, Kei 2, Kei 3, Kei 5, Kei 6, Log 1, Log 2,
Log 4, Per 2, Per 4, Red 2, Red 4, and Red 5 populations (Table 1). The
selected herbicides were glyphosate (Roundup
PowerMAX\textsuperscript \textregistered, Bayer Crop Science, Saint
Louis, MO, US) applied at 870 g ae ha\textsuperscript{-1} plus 2040 g
ha\textsuperscript{-1} ammonium sulfate (DSM Chemicals North America
Inc., Augusta, GA); fomesafen (Flexstar\textsuperscript \textregistered,
Syngenta Crop Protection, Greensboro, NC, USA) applied at 206 g ai
ha\textsuperscript{-1} plus 0.5 L ha\textsuperscript{-1} of non-ionic
surfactant (Induce\textsuperscript \textregistered, Helena
Agri-Enterprises, Collierville, TN, USA); and lactofen
(Cobra\textsuperscript \textregistered, Valent USA LLC Agricultural
Products, Walnut Creek, CA, USA) applied at 280 g ai
ha\textsuperscript{-1} plus 0.5 L ha\textsuperscript{-1} of non-ionic
surfactant.

Herbicide treatments were applied to 8-10 cm tall \emph{A. palmeri}
plants with a single-tip chamber sprayer (DeVries Manufacturing Corp.,
Hollandale, MN, USA). The sprayer had an 8001 E nozzle (Spraying Systems
Co., North Avenue, Wheaton, IL, US) calibrated to deliver 140 L
ha\textsuperscript{-1} spray volume at 135 kPa at a speed of 2.3 km
h\textsuperscript{-1}. \emph{A. palmeri} populations were visually
assessed 21 days after treatment (DAT) as dead or alive. Plants were
considered alive when prominent green tissue was observed in growing
plants, while dead plants were completely necrotic.

\hypertarget{statistical-analysis}{%
\subsection{Statistical Analysis}\label{statistical-analysis}}

\hypertarget{genotypic-and-phenotypic-validation-of-epsps--and-ppo-inhibitor-resistance-in-a.-palmeri}{%
\subsubsection{\texorpdfstring{Genotypic and Phenotypic validation of
EPSPS- and PPO-inhibitor resistance in \emph{A.
palmeri}}{Genotypic and Phenotypic validation of EPSPS- and PPO-inhibitor resistance in A. palmeri}}\label{genotypic-and-phenotypic-validation-of-epsps--and-ppo-inhibitor-resistance-in-a.-palmeri}}

The number of EPSPS- or PPO-inhibitor-resistant \emph{A. palmeri}
individuals in the genotypic assays was converted to a percentage scale:

\emph{Equation 1}: \[G=\frac{S}{T} * 100 \] \noindent where \emph{G} is
the \% EPSPS- or PPO-inhibitor-resistant \emph{A. palmeri} individuals,
\emph{S} is the total number of \emph{A. palmeri} individuals positive
for herbicide resistance, and \emph{T} is the total number of \emph{A.
palmeri} individuals (n=3 to 5) screened for herbicide resistance in
genotypic assays. Fomesafen and lactofen are PPO-inhibitor herbicides;
thus, \emph{G} is same for both.

The number of surviving \emph{A. palmeri} individuals in the phenotypic
assay were converted into a percentage scale:

\emph{Equation 2}: \[P=\frac{X}{T} * 100 \]

Where \emph{P} is the \% alive \emph{A. palmeri} individuals after
herbicide treatment in phenotypic assay (glyphosate, fomesafen, or
lactofen), \emph{X} is the total number of alive \emph{A. palmeri}
individuals 21 DAT and \emph{T} is the total number of \emph{A. palmeri}
individuals (n=40) treated with each herbicide. The P (\%) can be
determined only for 19 populations, as only 19 populations were screened
to herbicides. Data of two runs are combined.

The correlation between \emph{G} and \emph{P} for each herbicide
(glyphosate, fomesafen, and lactofen) and between the two PPO-inhibitor
herbicides, fomesafen and lactofen, were performed with Pearson's
analysis using the \emph{cor.test} function of R statistical software
version 3.6.1 (\emph{R} 2019). The correlation value varies from -1 and
1, where 1 is the total positive correlation, -1 the total negative
correlation, and 0 indicate no linear correlation. Pearson's analysis
tests the hypothesis that correlation between two variables is equal to
0 (null hypothesis). If \emph{P}-value \textgreater{} 0.05, the
probability is \textgreater{} 5\% that a correlation of some magnitude
between two variables could occur by chance alone assuming null
hypothesis is true; thus, no correaltion between variables. The \emph{G}
and \emph{P} correlation was performed with the 19 \emph{A. palmeri}
populations treated with herbicide in the phenotypic assay as well as
their respective genotypic assays results.

The individuals screened in genotypic and phenotypic assays are parent
and progenies, respectively. This methodology was chosen to simulate a
real farm scenario whereas growers could collect suspected herbicide
resistant leaf samples and mail them to a laboratory for faster
genotypic herbicide resistance confirmation.

\hypertarget{random-forest}{%
\subsubsection{Random Forest}\label{random-forest}}

Random forest is a powerful ensemble machine learning algorithm which
generates and combines multiple decision trees to reach a consensus. The
random forest procedure is described in detail by Breiman (Breiman 2001)
and Biau (Biau and Scornet 2016). In short, the random forest analysis
is largely based on two parameters: \emph{ntree}, which is the number of
decision trees, and \emph{mtry}, the number of different predictors
tested in each tree. For each decision tree, a subsample of observations
from the data are selected with replacement to train the trees
(bootstrap aggregating). These ``in-bag'' samples include approximately
66\% of the total data and some observations may be repeated in each new
training data set since this sampling is done with replacement. The
remaining 33\% of the data are designated ``Out-of-Bag'' or OOB samples
and are used in an internal cross-validation technique to estimate the
model performance error. To evaluate the importance of an explanatory
variable (or predictor), the random forest measures both the decrease in
model performance accuracy as calculated by the OOB error and the
decrease in the Gini Index value. The Gini Index value (mean decrease in
accuracy) is the mean of a total variable decrease of a node impurity,
weighted by the samples proportion reaching that node in each individual
decision tree. Therefore, variables with a large Gini Index value
indicates higher variable importance, and are more important for data
classification. Random Forest has been used to described the incidence
of crop disease (Langemeier et al. 2016) and glyphosate resistance in
\emph{Amaranthus} spp. (Vieira et al. 2018).

The random forest analysis was performed with the \emph{randomForest}
package of R statistical software to describe the influence of
\emph{EPSPS} gene amplification (genotypic), PPO-inhibitor resistance
(genotypic), location (county), agronomic practices (e.g., tillage,
irrigation, current and previous cropping-system), and weed demographics
(e.g., density and distribution) on EPSPS-inhibitor resistance in
\emph{A. palmeri} in southwestern Nebraska (Table 1). \emph{EPSPS} gene
copy number (genotypic) was included as an explanatory variable to test
the robustness of random forest as it is highly correlated to glyphosate
resistance in \emph{A. palmeri} (Gaines et al. 2019). The \emph{A.
palmeri} populations with at least one individual with \textgreater{} 2
\emph{EPSPS} copy number was considered resistant (Table 2); therefore,
EPSPS-inhibitor resistance was classified as Yes (\textgreater{} 2
\emph{EPSPS} copy number) or No.~For this analysis, the \emph{ntree}
parameter was set to 500, whereas \emph{mtry} was set to 2. Parameters
numbers selected are standard for the sample size of the present study.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{genotypic-and-phenotypic-validation-of-epsps--and-ppo-inhibitor-resistance-in-a.-palmeri-1}{%
\subsection{\texorpdfstring{Genotypic and phenotypic validation of
EPSPS- and PPO-inhibitor resistance in \emph{A.
palmeri}}{Genotypic and phenotypic validation of EPSPS- and PPO-inhibitor resistance in A. palmeri}}\label{genotypic-and-phenotypic-validation-of-epsps--and-ppo-inhibitor-resistance-in-a.-palmeri-1}}

The \emph{EPSPS} amplification was found in 63\% of the 51 \emph{A.
palmeri} populations analysed (Table 2). Phenotypic analysis of a subset
of these populations confirmed the genotypic analysis data, in that a
positive correlation (0.83; \emph{P}-value=0.0000) was observed between
\emph{G} and \emph{P} assays (Figure 1 and Table 3). Seven \emph{A.
palmeri} populations tested negative (\emph{G}=0) for glyphosate
resistance in the genotypic assay but three of these populations showed
low (\emph{P}=18\%, Hay 1), moderate (\emph{P}=35\%, Hay 4) and high
(\emph{P}=75\%, Red 5) survival after glyphosate treatment in the
phenotypic assays (Figure 1). The other four populations (Dun 3, Hay 3,
Log 2 and Kei 3) that tested negative (\emph{G}=0\%) for EPSPS-inhibitor
resistance in the genotypic assay showed less than 15\% glyphosate
survival (\emph{P}).

Fomesafen and lactofen resulted in less than 40\% survival of \emph{A.
palmeri} populations in the phenotypic assays (Figure 2). The
correlation between \emph{G} and \emph{P} for PPO-inhibitor resistance
in \emph{A. palmeri} populations was inconsistent (Table 3). While a
high \emph{G} and \emph{P} correlation (0.52; \emph{P}-value=0.0217) was
observed for fomesafen (Figure 2A), no \emph{G} and \emph{P} correlation
(-0.05; \emph{P}-value=0.84) was found for lactofen (Figure 2B).
\emph{A. palmeri} populations Dun 5, Kei 2, Kei 5, and Log 4 were
\emph{segregating for PPO resistance (Supp Table)} in the genotyping
assay but individuals in these populations were sensitive to lactofen
treatment (\emph{P}=0\%, Figure 2B). However, these populations showed
higher frequency of resistant individuals. For example, nearly 30\% of
individuals from the Log 4 population survived fomesafen treatment
(Figure 2A). In contrast, \emph{A. palmeri} populations Cha 3, Kei 6,
Per 2, and Red 5 tested negative for \(\triangle\)G210 or R128 het
mutations (\emph{G}=0\%) but over 15\% of populations survived both
fomesafen and lactofen treatment. Also, \emph{A. palmeri} populations
Kei 3, Per 4, and Dun 4 showed 38, 25, and 18\% survival after fomesafen
treatment but less than 15\% for lactofen. There was no correlation
(0.23; \emph{P}-value=0.34) between fomesafen and lactofen in the
phenotypic assay (Table 3).

According to the genotypic assay, nearly 70\% of the 51 \emph{A.
palmeri} populations from Southwestern Nebraska were confirmed resistant
to PPO-inhibitor herbicides. In the 34 PPO-inhibitor-resistant \emph{A.
palmeri} populations, 32 contain the \(\triangle\)G210 deletion in the
\emph{PPX2} gene, while the R128M/G endowed PPO-inhibitor resistance in
two \emph{A. palmeri} populations. Nearly 14\% of \emph{A. palmeri}
populations had all individuals resistant, 53\% were segregating for
resistance, and 33\% had no mutation detected to PPO-inhibitor
herbicides. In addition, based on \emph{EPSPS} gene amplification, our
study showed that 10\% of \emph{A. palmeri} populations had all
individuals resistant to glyphosate, 53\% were segregating for
resistance, and 37\% were susceptible to glyphosate (Table 2). Multiple
resistance (EPSPS- and PPO-inhibitor herbicides) was present in 41\% of
\emph{A. palmeri} populations of southwestern Nebraska (Table 2), while
6, 11, and 13 \emph{A. palmeri} populations were susceptible to both
herbicides, resistant to only glyphosate-, and resistant to only
PPO-inhibitor herbicides, respectively (Figure 3).

\hypertarget{random-forest-1}{%
\subsection{Random Forest}\label{random-forest-1}}

The final OOB error rate of the random forest analysis was 13.33\%,
meaning that \textgreater{} 86\% of OOB samples were adequately
classified by the model. The random forest analysis ranked (high to low)
\emph{EPSPS} gene amplification \textgreater{} county \textgreater{}
current crop \textgreater{} previous crop \textgreater{} \emph{A.
palmeri} density \textgreater{} tillage \textgreater{} irrigation
\textgreater{} \emph{A. palmeri} distribution \textgreater{}
PPO-inhibitor resistance as the factors influencing the presence of
EPSPS-inhibitor resistance in \emph{A. palmeri} of southwestern Nebraska
(Figure 4). The selection of \emph{EPSPS} gene amplification as the top
predictor shows the robustness of this model, since \emph{EPSPS} gene
amplification is known to confer resistance to glyphosate. County was
the second most important factor for the presence of EPSPS-inhibitor
resistant \emph{A. palmeri}. The presence of EPSPS-inhibitor resistance
in \emph{A. palmeri} was found in at least one population of all
surveyed counties from southwestern Nebraska (Table 2). The lowest
number of EPSPS-inhibitor-resistant populations was found at Hayes (Hay)
and Perkins (Per) County with 1 (out 5) and 2 (out 6), respectively.
Also, current and previous crops strongly influenced the presence of
EPSPS-inhibitor resistance in \emph{A. palmeri} populations. Five
\emph{A. palmeri} populations (Dun 4, Dun 5, Fro 1, Kei 2, and Lin 2)
had 100\% resistant individuals and these populations were all found in
current corn or soybean crops preceded by a corn or sorghum crops
(Figure 5A). In contrast, \emph{EPSPS} gene amplification did not occur
in 19 \emph{A. palmeri} populations, from which only two populations
were found in corn and soybean rotations (Fro 5 and Hit 5; Table 2). The
majority of EPSPS-inhibitor-susceptible \emph{A. palmeri} populations
were found in rotations of corn, soybean, sorghum, wheat, fallow, and
other crops (e.g., alfalfa, dry beans and field peas; Figure 5A).

Nearly 70\% of EPSPS-inhibitor-susceptible \emph{A. palmeri} populations
were resistant to PPO-inhibitor herbicides (Table 2). A similar trend of
crop diversity was observed for PPO-inhibitor resistance in \emph{A.
palmeri}, except a crop rotation of corn-sorghum-soybean was found in
locations with only PPO-inhibitor-resistant individuals (Figure 5B). The
random forest analysis was performed on EPSPS-inhibitor resistance only
due to the robustness of \emph{EPSPS} gene amplification as a positive
control to detect resistance in \emph{A. palmeri}, which is different
from our results with PPO-inhibitor resistance.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The high correlation between \emph{G} and \emph{P} for EPSPS-inhibitor
resistance demonstrates that most GR \emph{A. palmeri} populations from
southwestern Nebraska are resistant due to \emph{EPSPS} gene
amplification. However, other resistance mechanisms are likely present
since a few \emph{A. palmeri} populations showed no \emph{EPSPS} gene
amplification despite a majority of the individuals survived glyphosate
(870 g ae ha\textsuperscript{-1}) application in the phenotypic assay.
\emph{A. palmeri} was the first identified weed to evolve glyphosate
resistance via \emph{EPSPS} gene amplification (Gaines et al. 2010),
followed by \emph{Kochia scoparia}, \emph{Amaranthus tuberculatus},
\emph{Lolium perenne ssp. multiflorum}, \emph{Bromus diandrus},
\emph{Eleusine indica}, \emph{Chloris truncata}, and \emph{Amaranthus
hybridus} (Patterson et al. 2018, Sammons and Gaines 2014). The
\emph{EPSPS} gene amplification mechanism is widespread in \emph{A.
palmeri} (Gaines et al. 2019, Sammons and Gaines 2014), albeit other
EPSPS-inhibitor resistance mechanisms have arisen, including
\emph{Pro106} mutation in the \emph{EPSPS} gene and reduced glyphosate
absorption/translocation (Dominguez-Valenzuela et al. 2017,
Palma-Bautista et al. 2019, Sammons and Gaines 2014). It remains unknown
whether the \emph{A. palmeri} populations (e.g., Red 5) which have low
or no \emph{EPSPS} gene amplification harbor additional resistance
mechanisms, warranting further investigations.

Gene amplification is an important evolutionary mechanism enabling weeds
and other pests (Bass and Field 2011, Remnant et al. 2013) to evolve
resistance to pesticides. Research on \emph{EPSPS} gene amplification
molecular basis in weed species is underway but more work is still
needed to unveil this complex adaptative trait. The genetics of
\emph{EPSPS} gene amplification in weed species follows Mendelian
(\emph{K. scoparia}) (Jugulam et al. 2014) and non-Mendelian (\emph{A.
palmeri} (Giacomini et al. 2019) and \emph{B. diandrus} (Malone et al.
2016)) inheritance patterns. Over 100 \emph{EPSPS} gene copies have been
documented in \emph{A. palmeri} but maximum of only 10 have been
observed in \emph{K. scoparia} (Kumar et al. 2015, Wiersma et al. 2015).
The \emph{EPSPS} gene copy variation in \emph{A. palmeri} is a result of
the extrachromosomal circular DNA transmitted to the next generation by
tethering to mitotic and meiotic chromosomes (eccDNA) (Koo et al. 2018),
while in \emph{K. scoparia}, \emph{EPSPS} copies are arranged in tandem
repeats at a single locus (Patterson et al. 2018). Segregation for
\emph{EPSPS} copy number in \emph{A. palmeri} families (F\scriptsize 1
\normalsize and F\scriptsize 2 \normalsize) is transgressive, with
individuals varying in \emph{EPSPS} gene amplification levels even among
clonal plants (Giacomini et al. 2019). Transgressive segregation for
\emph{EPSPS} in \emph{A. palmeri} might explain the variable
\emph{EPSPS} copy numbers between individuals within populations
screened from southwestern Nebraska (Table 2). Gene amplification
coupled with the dioecious nature of \emph{A. palmeri} are valuable
traits for this weed that help increase its genetic complexity and allow
it to adapt to current US cropping systems.

According to the genotypic assay, most of \emph{A. palmeri} populations
from southwestern Nebraska also showed resistance to PPO-inhibitor
herbicides. The mechanism of resistance is largely due to the
\(\triangle\)G210 deletion but also due to R128M/G mutations in the
\emph{PPX2} gene (Table 2). It has been demonstrated that a mutated PPO
enzyme has reduced affinity for several PPO-inhibitor herbicides in
\emph{A. palmeri} (Schwartz-Lazaro et al. 2017). Application of
fomesafen (206 g ai ha\textsuperscript{-1}) and lactofen (280 g ai
ha\textsuperscript{-1}) provided high mortality in \emph{A. palmeri}
populations from southwestern Nebraska, including \emph{A. palmeri}
populations with 100\% individuals with \(\triangle\)G210 deletion.
Therefore, it is possible that most of sprayed \emph{A. palmeri}
individuals in the phenotypic assay had no mutation for PPO-inhibitor
resistance. High \emph{A. palmeri} mortality suggests that populations
are segregating for PPO-inhibitor resistance at a low level. Still,
there is correlation between \emph{G} and \emph{P} for fomesafen but not
lactofen, which was sprayed slightly above the labled rate of 218 g ai
ha\textsuperscript{-1}. The validation for PPO-inhibitor resistance
presented here is limited by the segregating nature of \emph{A. palmeri}
collected from field populations, since individuals in the phenotypic
assays were not the same as those screened in the genotypic assays. This
could be also explained in several ways, 1) the number of individuals
sampled for the \emph{G} assay study may have been too low for the
objective of validation; 2) high herbicide rate resulting in high
individual mortality; 3) greenhouse conditions were ideal and plants had
no stress during application of PPO-inhibitor herbicides, which might be
different from plants under field conditions of southwestern Nebraska
(Schafer et al. 2014); and 4) it is likely that plants were smaller than
usual because of small volume of cone-tainers, the size of a plant
strongly impacts the level of resistance, with smaller plants showing
less resistance than larger plants (Coburn 2017).

Populations of \emph{A. palmeri} have been reported to be resistant to
HPPD-(Jhala et al. 2014), PSII-(Jhala et al. 2014) and EPSPS-inhibitor
herbicides (Chahal et al. 2017) in Nebraska. In 2014, a survey with
\emph{Amaranthus} spp. in southwestern Nebraska showed widespread
EPSPS-inhibitor resistance for \emph{A. tuberculatus} (81\%) but not for
\emph{A. palmeri} (6\%) (Vieira et al. 2018). Our survey showed a high
number of EPSPS-inhibitor-resistant/segregating \emph{A. palmeri}
populations in southwestern Nebraska (\textgreater60\%). The rapid
evolution of EPSPS-inhibitor-resistant \emph{A. palmeri} in southwestern
Nebraska raised questions about whether populations were introduced via
seed/gene flow or arose independently. Although we did not test the
hypothesis, the random forest analysis did shed some light on
EPSPS-inhibitor resistance evolution in \emph{A. palmeri}. The random
forest analysis results suggested that \emph{EPSPS} gene amplification,
county, and current and previous crops were the top factors influencing
GR \emph{A. palmeri} in southwestern Nebraska. The \emph{EPSPS} gene
amplification as the primary factor for glyphosate resistance in
\emph{A. palmeri} showed the robustness of random forest analysis, while
county influence on \emph{EPSPS}-inhibitor resistance in \emph{A.
palmeri} highlighted how practices for delaying EPSPS-inhibitor
resistance evolution should start at the landscape level. It is likely
that growers with common goals may implement best weed management
tactics to delay EPSPS-inhibitor resistance evolution and prevent
seed/gene flow of herbicide resistant weeds. Nonetheless, it is striking
that PPO-inhibitor resistance was common in counties with low/no
EPSPS-inhibitor resistance. For instance, most
EPSPS-inhibitor-susceptible \emph{A. palmeri} were found in Hayes County
in areas of sorghum or corn in rotation with wheat (Table 1), which are
less glyphosate-dependent crops. Although best weed management practices
at the landscape level are encouraged, our results suggest that growers
were not necessarily implementing good management practices but rather,
using less EPSPS-inhibitor herbicide.

The high presence of \emph{A. palmeri} populations with 100\% resistance
to EPSPS-inhibitors in less diverse cropping systems (e.g., corn-soybean
rotation) suggests the influence of repeated glyphosate application on
\emph{A. palmeri} evolution to EPSPS-inhibitor resistance. The
occurrence of EPSPS-inhibitor-resistant \emph{A. palmeri} populations is
reduced in rotations with more diversified crops (Figure 5A), most
likely due to rotations with glyphosate sensitive crops (no/less
glyphosate use). For example, according to a survey, 2,4-D and
metsulfuron-methyl are the most used POST herbicides in grain sorghum
and wheat in Nebraska, respectively (Sarangi and Jhala 2018). Crop
diversity exerts a different selection pressure on weed communities,
including canopy closure timing, seeding, and harvest date, which aid on
reducing the dominance of single weed species (Andrade et al. 2017).
Nonetheless, it is likely that Nebraska growers still rely on other
herbicide sites of action for weed management, as most
EPSPS-inhibitor-susceptible \emph{A. palmeri} populations were resistant
to PPO-inhibitor herbicides (e.g., fomesafen). Similar to
EPSPS-inhibitor resistance, the PPO-inhibitor-resistant \emph{A.
palmeri} was also found more often in less diverse crop rotations
(Figure 5B). Despite not having long-term herbicide application records
for the surveyed area, it has been demonstrated that overreliance on a
single or few herbicide SOA in areas with low crop diversity contributed
to resistance (Hicks et al. 2018). In addition, it has been shown that
herbicide mixture (multiple SOA in one application) is more effective
for delaying herbicide weed resistance than herbicide rotation (multiple
applications, each with a single SOA) (Beckie and Reboud 2009, Evans et
al. 2016). Thus, without herbicide mixture, increased crop diversity is
not enough to reduce the risk of herbicide resistance evolution in weed
species.

An emerging concern in weed science is the ability of some species to
stack genes for multiple herbicide resistance in a single population.
Five- and six-way herbicide resistance has been reported in an \emph{A.
tuberculatus} population from Illinois (Strom et al. 2019) and Missouri
(Shergill et al. 2018), respectively. The 19 \emph{A. palmeri}
populations used in the whole-plant phenotypic assay also tested
positive (\textgreater80\% of individuals) to an ALS-inhibitor herbicide
(imazethapyr {[}70 g ai ha\textsuperscript{-1}{]}; data not shown).
Moreover, nearly half of the \emph{A. palmeri} populations in this study
possessed multiple herbicide resistance (EPSPS and PPO; Figure 3). Thus,
it is likely that two- and three-way resistance exists in most \emph{A.
palmeri} populations from southwestern Nebraska. The random forest
analysis suggested no/low influence of PPO-inhibitor resistance on
EPSPS-inhibitor resistance, indicating \emph{EPSPS} gene amplification
and PPO-inhibitor resistance mutations (\(\triangle\)G210 or R128M/G)
are not genetically linked. Cosegregation of distinct SOA resistance
mechanisms is unlikely but not impossible as it has been shown that ALS-
and PPO-inhibitor resistance are genetically linked in a \emph{A.
tuberculatus} population (Tranel et al. 2017).

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Herein we demonstrate the widespread occurrence of EPSPS- and
PPO-inhibitor-resistant \emph{A. palmeri} in southwestern Nebraska.
Glyphosate resistance via \emph{EPSPS} gene amplification was present in
the majority of \emph{A. palmeri} populations from Nebraska, supporting
the use of genotypic assays for faster detection of
\emph{EPSPS}-inhibitor resistance in this region. Novel or non-target
glyphosate resistance mechanisms may also be present in few \emph{A.
palmeri} populations, but at a relatively low frequency. PPO-inhibitor
resistance was also present in these populations, but was less
correlated with known PPO-inhibitor resistance mutations, warranting the
use of both genotypic and whole-plant bioassays for testing
PPO-inhibitor resistance. The EPSPS- or PPO-inhibitor-resistant \emph{A.
palmeri} populations were found in areas of low crop diversity,
suggesting resistance evolution was mostly due to high selection
pressures of EPSPS- and PPO-inhibitor herbicides. Great progress has
been made towards understanding the \emph{A. palmeri} molecular basis of
resistance, but the continuous spread of herbicide resistance to new
geographies is evident. Thus, there is a need to change the current
methods for crop and weed management, including dependence on
corn/soybean rotation, use herbicide mixtures with different SOA rather
than sequential applications, and bring non-herbicide weed management
innovations to US cropping systems.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

This research received no specific grant from any funding agency. We
also appreciate the help of Victor Ribeiro's assistance with the
greenhouse projects.

\hypertarget{conflicts-of-interest}{%
\section{Conflicts of Interest}\label{conflicts-of-interest}}

The authors declare no conflict of interest.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-andrade_weed_2017}{}%
Andrade JF, Satorre EH, Ermácora CM, Poggio SL (2017) Weed communities
respond to changes in the diversity of crop sequence composition and
double cropping. Weed Research 57:148--158

\leavevmode\hypertarget{ref-bass_gene_2011}{}%
Bass C, Field LM (2011) Gene amplification and insecticide resistance.
Pest Management Science 67:886--890

\leavevmode\hypertarget{ref-beckie_selecting_2009}{}%
Beckie HJ, Reboud X (2009) Selecting for weed resistance: Herbicide
rotation and mixture. Weed Technology 23:363--370

\leavevmode\hypertarget{ref-biau_random_2016}{}%
Biau G, Scornet E (2016) A random forest guided tour. TEST 25:197--227

\leavevmode\hypertarget{ref-breiman_random_2001}{}%
Breiman L (2001) Random Forests. Machine Learning 45:5--32

\leavevmode\hypertarget{ref-chahal_glyphosate-resistant_2017}{}%
Chahal PS, Varanasi VK, Jugulam M, Jhala AJ (2017) Glyphosate-resistant
Palmer amaranth (\emph{Amaranthus} \emph{Palmeri}) in Nebraska:
Confirmation, EPSPS gene amplification, and response to POST corn and
soybean herbicides. Weed Technology 31:80--93

\leavevmode\hypertarget{ref-chatham_multistate_2015}{}%
Chatham LA, Bradley KW, Kruger GR, Martin JR, Owen MDK, Peterson DE,
Mithila J, Tranel PJ (2015) A Multistate study of the association
between glyphosate resistance and EPSPS gene amplification in waterhemp
(\emph{Amaranthus} \emph{Tuberculatus}). Weed Science 63:569--577

\leavevmode\hypertarget{ref-coburn_influence_2017}{}%
Coburn C (2017) Influence of Experimental Methods on Herbicide
Resistance Confirmation. Ph.D dissertation. Laramie, WY: University of
Wyoming. 145 p

\leavevmode\hypertarget{ref-culpepper_glyphosate-resistant_2006}{}%
Culpepper AS, Grey TL, Vencill WK, Kichler JM, Webster TM, Brown SM,
York AC, Davis JW, Hanna WW (2006) Glyphosate-resistant Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) confirmed in Georgia. Weed Science
54:620--626

\leavevmode\hypertarget{ref-dominguez-valenzuela_first_2017}{}%
Dominguez-Valenzuela JA, Gherekhloo J, Fernández-Moreno PT,
Cruz-Hipolito HE, Alcántara-de la Cruz R, Sánchez-González E, De Prado R
(2017) First confirmation and characterization of target and non-target
site resistance to glyphosate in Palmer amaranth (\emph{Amaranthus}
\emph{Palmeri}) from Mexico. Plant Physiology and Biochemistry
115:212--218

\leavevmode\hypertarget{ref-doyle_rapid_1987}{}%
Doyle JJ, Doyle JL (1987) A rapid DNA isolation procedure for small
quantities of fresh leaf tissue. Phytochemical Bull 19:11--15

\leavevmode\hypertarget{ref-evans_managing_2016}{}%
Evans JA, Tranel PJ, Hager AG, Schutte B, Wu C, Chatham LA, Davis AS
(2016) Managing the evolution of herbicide resistance: Managing the
evolution of herbicide resistance. Pest Manag Sci 72:74--80

\leavevmode\hypertarget{ref-farmer_evaluating_2017}{}%
Farmer JA, Webb EB, Pierce RA, Bradley KW (2017) Evaluating the
potential for weed seed dispersal based on waterfowl consumption and
seed viability. Pest Management Science 73:2592--2603

\leavevmode\hypertarget{ref-gaines_molecular_2019}{}%
Gaines TA, Patterson EL, Neve P (2019) Molecular mechanisms of adaptive
evolution revealed by global selection for glyphosate resistance. New
Phytologist 0

\leavevmode\hypertarget{ref-gaines_mechanism_2011}{}%
Gaines TA, Shaner DL, Ward SM, Leach JE, Preston C, Westra P (2011)
Mechanism of Resistance of Evolved Glyphosate-Resistant Palmer Amaranth
(\emph{Amaranthus} \emph{Palmeri}). J Agric Food Chem 59:5886--5889

\leavevmode\hypertarget{ref-gaines_interspecific_2012}{}%
Gaines TA, Ward SM, Bukun B, Preston C, Leach JE, Westra P (2012)
Interspecific hybridization transfers a previously unknown glyphosate
resistance mechanism in \emph{Amaranthus} species. Evolutionary
Applications 5:29--38

\leavevmode\hypertarget{ref-gaines_gene_2010}{}%
Gaines TA, Zhang W, Wang D, Bukun B, Chisholm ST, Shaner DL, Nissen SJ,
Patzoldt WL, Tranel PJ, Culpepper AS, Grey TL, Webster TM, Vencill WK,
Sammons RD, Jiang J, Preston C, Leach JE, Westra P (2010) Gene
amplification confers glyphosate resistance in \emph{Amaranthus}
\emph{Palmeri}. Proceedings of the National Academy of Sciences
107:1029--1034

\leavevmode\hypertarget{ref-giacomini_two_2017-1}{}%
Giacomini DA, Umphres AM, Nie H, Mueller TC, Steckel LE, Young BG, Scott
RC, Tranel PJ (2017) Two new PPX2 mutations associated with resistance
to PPO-inhibiting herbicides in \emph{Amaranthus} \emph{Palmeri}. Pest
Management Science 73:1559--1563

\leavevmode\hypertarget{ref-giacomini_variable_2019}{}%
Giacomini DA, Westra P, Ward SM (2019) Variable Inheritance of Amplified
EPSPS Gene Copies in Glyphosate-Resistant Palmer Amaranth
(\emph{Amaranthus} \emph{Palmeri}). Weed Science 67:176--182

\leavevmode\hypertarget{ref-gossett_resistance_1992}{}%
Gossett BJ, Murdock EC, Toler JE (1992) Resistance of Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) to the dinitroaniline herbicides.
Weed Technology 6:587--591

\leavevmode\hypertarget{ref-hamilton_weeds_1958}{}%
Hamilton KC, Arle HF (1958) Weeds od crops in Southern Arizona. Tucson,
AZ: College of Agriculture, University of Arizona. 66 p

\leavevmode\hypertarget{ref-heap_list_2019-1}{}%
Heap I (2019) List of Herbicide Resistant Weeds by Weed Species.
\url{http://www.weedscience.org/Summary/Species.aspx}. Accessed June 12,
2019

\leavevmode\hypertarget{ref-hicks_factors_2018}{}%
Hicks HL, Comont D, Coutts SR, Crook L, Hull R, Norris K, Neve P, Childs
DZ, Freckleton RP (2018) The factors driving evolved herbicide
resistance at a national scale. Nature Ecology \& Evolution 2:529--536

\leavevmode\hypertarget{ref-horak_biotypes_1995}{}%
Horak MJ, Peterson DE (1995) Biotypes of Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) and common waterhemp
(\emph{Amaranthus} \emph{Rudis}) are resistant to imazethapyr and
thifensulfuron. Weed Technology 9:192--195

\leavevmode\hypertarget{ref-jhala_confirmation_2014}{}%
Jhala AJ, Sandell LD, Rana N, Kruger GR, Knezevic SZ (2014) Confirmation
and control of triazine and 4-hydroxyphenylpyruvate
dioxygenase-inhibiting herbicide-resistant Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) in Nebraska. Weed Technology
28:28--38

\leavevmode\hypertarget{ref-jugulam_tandem_2014}{}%
Jugulam M, Niehues K, Godar AS, Koo D-H, Danilova T, Friebe B, Sehgal S,
Varanasi VK, Wiersma A, Westra P, Stahlman PW, Gill BS (2014) Tandem
amplification of a chromosomal segment harboring
5-enolpyruvylshikimate-3-phosphate synthase locus confers glyphosate
resistance in \emph{Kochia} \emph{Scoparia}. Plant Physiology
166:1200--1207

\leavevmode\hypertarget{ref-kohrt_confirmation_2017}{}%
Kohrt JR, Sprague CL, Nadakuduti SS, Douches D (2017) Confirmation of a
Three-Way (Glyphosate, ALS, and Atrazine) Herbicide-Resistant Population
of Palmer Amaranth (Amaranthus palmeri) in Michigan. Weed Science
65:327--338

\leavevmode\hypertarget{ref-koo_extrachromosomal_2018}{}%
Koo D-H, Molin WT, Saski CA, Jiang J, Putta K, Jugulam M, Friebe B, Gill
BS (2018) Extrachromosomal circular DNA-based amplification and
transmission of herbicide resistance in crop weed \emph{Amaranthus}
\emph{Palmeri}. Proc Natl Acad Sci USA 115:3332--3337

\leavevmode\hypertarget{ref-kumar_molecular_2015}{}%
Kumar V, Jha P, Giacomini D, Westra EP, Westra P (2015) Molecular basis
of evolved resistance to glyphosate and acetolactate synthase-inhibitor
herbicides in Kochia (\emph{Kochia} \emph{Scoparia}) accessions from
Montana. Weed sci 63:758--769

\leavevmode\hypertarget{ref-kumar_confirmation_2019}{}%
Kumar V, Liu R, Boyer G, Stahlman PW (2019) Confirmation of 2,4-D
resistance and identification of multiple resistance in a Kansas Palmer
amaranth (\emph{Amaranthus} \emph{Palmeri}) population. Pest Management
Science 0

\leavevmode\hypertarget{ref-kupper_population_2018}{}%
Küpper A, Manmathan HK, Giacomini D, Patterson EL, McCloskey WB, Gaines
TA (2018) Population genetic structure in glyphosate-resistant and
-susceptible Palmer amaranth (\emph{Amaranthus} \emph{Palmeri})
populations using genotyping-by-sequencing (GBS). Front Plant Sci 9

\leavevmode\hypertarget{ref-langemeier_factors_2016}{}%
Langemeier CB, Robertson AE, Wang D, Jackson-Ziems TA, Kruger GR (2016)
Factors affecting the development and severity of goss's bacterial wilt
and leaf blight of corn, caused by \emph{Clavibacter}
\emph{Michiganensis} subsp. \emph{Nebraskensis}. Plant Disease
101:54--61

\leavevmode\hypertarget{ref-malone_epsps_2016}{}%
Malone JM, Morran S, Shirley N, Boutsalis P, Preston C (2016) EPSPS gene
amplification in glyphosate-resistant \emph{Bromus} \emph{Diandrus}.
Pest Management Science 72:81--88

\leavevmode\hypertarget{ref-massinga_water_2003}{}%
Massinga RA, Currie RS, Trooien TP (2003) Water use and light
interception under Palmer amaranth (\emph{Amaranthus} \emph{Palmeri})
and corn competition. Weed Science 51:523--531

\leavevmode\hypertarget{ref-morgan_competitive_2001}{}%
Morgan GD, Baumann PA, Chandler JM (2001) Competitive impact of Palmer
amaranth (\emph{Amaranthus} \emph{Palmeri}) on cotton (\emph{Gossypium}
\emph{Hirsutum}) development and yield. Weed Technology 15:408--412

\leavevmode\hypertarget{ref-norsworthy_-field_2014}{}%
Norsworthy JK, Griffith G, Griffin T, Bagavathiannan M, Gbur EE (2014)
In-field movement of glyphosate-resistant Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) and its impact on cotton lint Yield:
Evidence supporting a zero-threshold strategy. Weed Science 62:237--249

\leavevmode\hypertarget{ref-oliveira_interspecific_2018}{}%
Oliveira MC, Gaines TA, Patterson EL, Jhala AJ, Irmak S, Amundsen K,
Knezevic SZ (2018) Interspecific and intraspecific transference of
metabolism-based mesotrione resistance in dioecious weedy
\emph{Amaranthus}. The Plant Journal 96:1051--1063

\leavevmode\hypertarget{ref-palma-bautista_reduced_2019}{}%
Palma-Bautista C, Torra J, Garcia MJ, Bracamonte E, Rojano-Delgado AM,
Alcántara-de la Cruz R, De Prado R (2019) Reduced Absorption and
Impaired Translocation Endows Glyphosate Resistance in Amaranthus
palmeri Harvested in Glyphosate-Resistant Soybean from Argentina. J
Agric Food Chem 67:1052--1060

\leavevmode\hypertarget{ref-patterson_glyphosate_2018}{}%
Patterson EL, Pettinga DJ, Ravet K, Neve P, Gaines TA (2018) Glyphosate
resistance and EPSPS gene duplication: Convergent evolution in multiple
plant species. J Hered 109:117--125

\leavevmode\hypertarget{ref-powles_evolved_2008}{}%
Powles SB (2008) Evolved glyphosate-resistant weeds around the world:
Lessons to be learnt. Pest Management Science 64:360--365

\leavevmode\hypertarget{ref-price_glyphosate-resistant_2011}{}%
Price AJ, Balkcom KS, Culpepper SA, Kelton JA, Nichols RL, Schomberg H
(2011) Glyphosate-resistant Palmer amaranth: A threat to conservation
tillage. Journal of Soil and Water Conservation 66:265--275

\leavevmode\hypertarget{ref-r_core_team_r:_2019}{}%
R: A Language and Environment for Statistical Computing (2019). Vienna,
Austria: R Foundation for Statistical Computing

\leavevmode\hypertarget{ref-rangani_novel_2019}{}%
Rangani G, Salas-Perez RA, Aponte RA, Knapp M, Craig IR, Mietzner T,
Langaro AC, Noguera MM, Porri A, Roma-Burgos N (2019) A Novel
Single-Site Mutation in the Catalytic Domain of Protoporphyrinogen
Oxidase IX (PPO) Confers Resistance to PPO-Inhibiting Herbicides. Front
Plant Sci 10

\leavevmode\hypertarget{ref-remnant_gene_2013}{}%
Remnant EJ, Good RT, Schmidt JM, Lumb C, Robin C, Daborn PJ, Batterham P
(2013) Gene duplication in the major insecticide target site, Rdl, in
\emph{Drosophila} \emph{Melanogaster}. Proc Natl Acad Sci USA
110:14705--14710

\leavevmode\hypertarget{ref-salas_resistance_2016}{}%
Salas RA, Burgos NR, Tranel PJ, Singh S, Glasgow L, Scott RC, Nichols RL
(2016) Resistance to PPO-inhibiting herbicide in Palmer amaranth from
Arkansas. Pest Management Science 72:864--869

\leavevmode\hypertarget{ref-salas-perez_frequency_2017}{}%
Salas-Perez RA, Burgos NR, Rangani G, Singh S, Refatti JP, Piveta L,
Tranel PJ, Mauromoustakos A, Scott RC (2017) Frequency of gly-210
deletion mutation among protoporphyrinogen oxidase inhibitor--resistant
Palmer amaranth (\emph{Amaranthus} \emph{Palmeri}) populations. Weed
Science 65:718--731

\leavevmode\hypertarget{ref-sammons_glyphosate_2014}{}%
Sammons RD, Gaines TA (2014) Glyphosate resistance: State of knowledge.
Pest Management Science 70:1367--1377

\leavevmode\hypertarget{ref-sarangi_statewide_2018-1}{}%
Sarangi D, Jhala AJ (2018) A Statewide Survey of Stakeholders to Assess
the Problem Weeds and Weed Management Practices in Nebraska. Weed
Technology 32:642--655

\leavevmode\hypertarget{ref-sauer_revision_1955}{}%
Sauer J (1955) Revision of the dioecious Amaranths. Madroño 13:5--46

\leavevmode\hypertarget{ref-sauer_recent_1957}{}%
Sauer J (1957) Recent migration and evolution of the dioecious
Amaranths. Evolution 11:11--31

\leavevmode\hypertarget{ref-sauer_dioecious_1972}{}%
Sauer JD (1972) The dioecious Amaranths: A new species name and major
range extensions. Madroño 21:426--434

\leavevmode\hypertarget{ref-schafer_rhizosphere_2014}{}%
Schafer JR, Hallett SG, Johnson WG (2014) Rhizosphere Microbial
Community Dynamics in Glyphosate-Treated Susceptible and Resistant
Biotypes of Giant Ragweed (Ambrosia trifida). Weed Science 62:370--381

\leavevmode\hypertarget{ref-schwartz-lazaro_resistance_2017}{}%
Schwartz-Lazaro LM, Norsworthy JK, Scott RC, Barber LT (2017) Resistance
of two Arkansas Palmer amaranth populations to multiple herbicide sites
of action. Crop Protection 96:158--163

\leavevmode\hypertarget{ref-shergill_investigations_2018}{}%
Shergill LS, Barlow BR, Bish MD, Bradley KW (2018) Investigations of
2,4-D and multiple herbicide resistance in a Missouri waterhemp
(\emph{Amaranthus} \emph{Tuberculatus}) population. Weed Science
66:386--394

\leavevmode\hypertarget{ref-smith_fodder_1900}{}%
Smith J (1900) Fodder and forage plants. Exclusive of the grasses.
Washington. 65 p

\leavevmode\hypertarget{ref-strom_characterization_2019}{}%
Strom SA, Gonzini LC, Mitsdarfer C, Davis AS, Riechers DE, Hager AG
(2019) Characterization of multiple herbicide--resistant waterhemp
(\emph{Amaranthus} \emph{Tuberculatus}) populations from Illinois to
VLCFA-inhibiting herbicides. Weed Science 67:369--379

\leavevmode\hypertarget{ref-tranel_target-site_2017}{}%
Tranel PJ, Wu C, Sadeque A (2017) Target-site resistances to ALS and PPO
inhibitors are linked in waterhemp (\emph{Amaranthus}
\emph{Tuberculatus}). Weed Science 65:4--8

\leavevmode\hypertarget{ref-varanasi_confirmation_2018}{}%
Varanasi VK, Brabham C, Norsworthy JK (2018a) Confirmation and
Characterization of Non--target site Resistance to Fomesafen in Palmer
amaranth (Amaranthus palmeri). Weed Science 66:702--709

\leavevmode\hypertarget{ref-varanasi_statewide_2018}{}%
Varanasi VK, Brabham C, Norsworthy JK, Nie H, Young BG, Houston M,
Barber T, Scott RC (2018b) A Statewide survey of PPO-inhibitor
resistance and the prevalent target-site mechanisms in Palmer amaranth
(\emph{Amaranthus} \emph{Palmeri}) accessions from Arkansas. Weed
Science 66:149--158

\leavevmode\hypertarget{ref-vieira_distribution_2018}{}%
Vieira BC, Samuelson SL, Alves GS, Gaines TA, Werle R, Kruger GR (2018)
Distribution of glyphosate-resistant \emph{Amaranthus} spp. In Nebraska.
Pest Management Science 74:2316--2324

\leavevmode\hypertarget{ref-ward_palmer_2013}{}%
Ward SM, Webster TM, Steckel LE (2013) Palmer Amaranth
(\emph{Amaranthus} \emph{Palmeri}): A Review. Weed Technology 27:12--27

\leavevmode\hypertarget{ref-wiersma_gene_2015}{}%
Wiersma AT, Gaines TA, Preston C, Hamilton JP, Giacomini D, Robin Buell
C, Leach JE, Westra P (2015) Gene amplification of
5-enol-pyruvylshikimate-3-phosphate synthase in glyphosate-resistant
\emph{Kochia} \emph{Scoparia}. Planta 241:463--474

\leavevmode\hypertarget{ref-wuerffel_distribution_2015}{}%
Wuerffel RJ, Young JM, Lee RM, Tranel PJ, Lightfoot DA, Young BG (2015)
Distribution of the G210 protoporphyrinogen oxidase mutation in Illinois
waterhemp (\emph{Amaranthus} \emph{Tuberculatus}) and an improved
molecular method for detection. Weed Science 63:839--845

\end{document}
